---
title: 'Homework #5'
author: "Helle Leth Skjetne"
date: "9/23/2020"
output: github_document
---

# Setting up rquired packages and library
```{r setup, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("nycflights13") #download the package if it is not already installed
library(nycflights13)
library(tidyverse)
```

# Split-apply-combine

In this part I will answer questions 1 - 5 related to Homework #5. The code will be included in the code chunk below:

## 1. Use ?nycflights13::flights to see the documentation of flights
```{r q1, echo=TRUE}
?nycflights13::flights
```

## 2. Use the glimpse() function to get a general sense of the flights data frame. How many rows and columns does flights have?
```{r q2, echo=TRUE}
flights <- nycflights13::flights #define the variable flights
glimpse(flights) #get a glimpse of the data file flights
###Answer:
  #Rows: 336,776
  #Columns: 19
```


## 3. Use filter() to create a data frame of just the flights that departed 2 hours late or more. What fraction of flights departed at least 2 hours late?
```{r q3, echo=TRUE}
### Create filter:
delayedflights <- filter(.data = flights, flights$dep_delay >= 2) #filter data frame for flights that departed ≥2 late
delayedflights
### Determine fraction of flights that were at least 2 hours late:
glimpse(delayedflights) #Rows: 120,382
glimpse(flights) #Rows: 336,776
fraction_late <- 120382/336776
fraction_late #a fraction of 0.3574542 flights were 2 hours or more late
```

## 4. Use mutate() and paste0() to create a column that contains the combined airline and flight number (e.g., “UA5041”).
```{r q4, echo=TRUE}
airline_flightnum <- mutate(flights, flightnum = paste0(flights$carrier, flights$flight))
airline_flightnum #contains the column "flightnum" with information about combined airline and flight number
```

## 5. Use group_by(), summarise(), and n() to make a data frame showing how many flights each aircraft (tail_num()) flew for each carrier throughout the whole data set.
```{r q5, echo=TRUE}
### Code to show number of flights each aircraft flew for each carrier:
num_flights <- flights %>%
  group_by(carrier, tailnum) %>%
  summarise(
    NumFlights = n()
  ) #"Warning": `summarise()` regrouping output by 'carrier' (override with `.groups` argument)
num_flights
```


