---
title: "Untitled"
author: "Helle Leth Skjetne"
date: "10/27/2020"
output: github_document
---

```{r setup, include=TRUE}
#load packages
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(ggpubr)
library(ggpmisc)

#load data
diamonds

# Set working directory
setwd("~/Documents/GEOL590/reproducible_data_analysis/homework/hw_10") #set working directory
getwd() #check working directory
```

# Linear Models

## Using the diamonds dataset, create a linear model of price as a function of the weight of the diamond (carat). According to that model, what price would you expect for a 1-carat diamond?
### Make some plots of the raw data, and of the model fits, to make an argument as to which model is more useful. 

```{r q1, include=TRUE}
carat_price <- lm(price ~ carat, data = diamonds)
summary(carat_price)

m2 <- update(carat_price, ~ . + carat)

thisDiamond <- data.frame(carat = 1)
(modelEstimate <- predict(m2, newdata = thisDiamond,
                          interval = "prediction", level = 0.95))
# the predicted price is $5500 of a 1 carat diamond

formula1 <- y ~ x

#create plot for fun
pricecarat <- ggplot(aes(x = carat, y = price), data = diamonds) +
  geom_point(fill=I("red"), color=I("black"), shape=21) +
  scale_x_continuous(lim = c(0, quantile(diamonds$carat, 0.99)) ) +
  scale_y_continuous(lim = c(0, quantile(diamonds$price, 0.99)) ) +
  ggtitle("Diamonds: Price vs. Carat") + 
  geom_smooth(method = 'lm') +
  stat_poly_eq(aes(label = paste0("atop(", ..eq.label.., ",", ..rr.label.., ")")), 
               formula = formula1, 
               parse = TRUE)
pricecarat
```


## Now create a linear model of price as a function of carat and color.
### What difference does it make if you convert color to a character vector before you make the model? Note that this is probably what you want to do in general, unless you really know what you’re doing.
### Make some plots of the raw data, and of the model fits, to make an argument as to which model is more useful. 
    
```{r q2, include=TRUE}
#linear model of price as a function of carat and color
carat_price_color <- lm(price ~ carat + color, data = diamonds)
carat_price_color
summary(carat_price_color)

# price = m1 * carat + m2 + color + b

glimpse(diamonds)
#levels)diamonds$color)
summary(mod) #carat_price_color

mod_unordered <- lm(price ~ carat + factor(color, ordered = FALSE), 
               data = diamonds)
summary(mod_unordered)

mod_char <- lm(price ~ carat + as.character(color), #a character is unordered, same values in Estimate for the one example above
               data = diamonds)


#a linear model of price as a function of carat and color
ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1) +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = "Color", 
                                          reverse = F,
                                          override.aes = list(alpha = 1, 
                                                              size = 2))) +  
  ggtitle('Price of Diamonds as a Function of Carat and Color') + labs(x = "Carat", y = "Price ($)") + 
  stat_poly_eq(aes(label = paste0("atop(", ..eq.label.., ",", ..rr.label.., ")")), 
               formula = formula1, 
               parse = TRUE) +
    geom_smooth(method = 'lm') +
  facet_wrap(vars(color))
```

# Objects and Methods

## Write a function that returns some kind of structured result, which you define as a new S3 class. Your class doesn’t need to be anything fancy, but it should be different from existing classes.

```{r q3, include=FALSE}

dog <- function(barks = FALSE, breed = "canine") {
  info <- list(
    noisy = barks,
    type = breed
  )
  
  ## Set the name for the class
  class(info) <- append(class(info), "dog")
  return(info)
}

#create a new object of this class by calling the function:
  ##for australian shepherd
  aussie <- dog()
  aussie
  aussie$noisy
  aussie$type
  
  ##for corgi
  corgi <- dog(barks = TRUE, breed = "corgi")
  corgi
  
  
NorthAmerican <- function(eatsBreakfast=TRUE,myFavorite="cereal")
{

        me <- list(
                hasBreakfast = eatsBreakfast,
                favoriteBreakfast = myFavorite
       )

        ## Set the name for the class
        class(me) <- append(class(me),"NorthAmerican")
        return(me)
}

```

## Write print and summarise methods for this class. These methods don’t need to be particularly useful, but they should be different from the default methods.

```{r q4a, include=TRUE}
setNoisy <- function(elObjecto, newValue) 
  {
  print("Calling the base setNoisy function")
  UseMethod("setNoisy", elObjecto)
  print("This has not yet been executed!")
}

setNoisy.default <- function(elObjecto, newValue)
{
  print("Cannot handle object.")
  return(elObjecto)
}

setNoisy.dog <- function(elObjecto, newValue)
{
  print("In setNoisy.dog and setting the value.")
  elObjecto$Noisy <- newValue
  return(elObjecto)
}

#test method above:
aussie <- dog()
aussie$noisy
aussie <- setNoisy(aussie, TRUE) #change "noisy" from FALSE to TRUE
  #[1] "Calling the base setNoisy function"
  #[1] "In setNoisy.dog and setting the value."
aussie <- setNoisy(aussie, "Are the cutest dogs even if they sometimes bark!")
  #[1] "Calling the base setNoisy function"
  #[1] "In setNoisy.dog and setting the value."
aussie$Noisy
  #[1] "Are the cutest dogs even if they sometimes bark!"

#use predefined accessors to get and set values held by an object:
getNoisy <- function(elObjecto)
{
  print("Calling the base getNoisy function")
  UseMethod("getNoisy", elObjecto)
  print("This has not been executed yet!")
}

getNoisy.default <- function(elObjecto)
{
  print("Error. This object cannot be handeled.")
  return(NULL)
}

getNoisy.dog <- function(elObjecto)
{
  print("In getNoisy.dog and returning the value")
  return(elObjecto$Noisy)
}

```

```{r qba, include=TRUE}
newfuction <- function(h) {
  for(i in 1:h) {
    j <- 1+i^2
    print(j)
  }
}
newfuction(9)
```