---
title: "Untitled"
author: "Helle Leth Skjetne"
date: "10/27/2020"
output: html_document
---

```{r setup, include=TRUE}
#load packages
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(ggpubr)
library(ggpmisc)

#load data
diamonds

# Set working directory
setwd("~/Documents/GEOL590/reproducible_data_analysis/homework/hw_10") #set working directory
getwd() #check working directory
```

# Linear Models

## Using the diamonds dataset, create a linear model of price as a function of the weight of the diamond (carat). According to that model, what price would you expect for a 1-carat diamond?
### Make some plots of the raw data, and of the model fits, to make an argument as to which model is more useful. 

```{r q1, include=TRUE}
carat_price <- lm(price ~ carat, data = diamonds)
summary(carat_price)

m2 <- update(carat_price, ~ . + carat)

thisDiamond <- data.frame(carat = 1)
(modelEstimate <- predict(m2, newdata = thisDiamond,
                          interval = "prediction", level = 0.95))
# the predicted price is $5500 of a 1 carat diamond

formula1 <- y ~ x

#create plot for fun
pricecarat <- ggplot(aes(x = carat, y = price), data = diamonds) +
  geom_point(fill=I("red"), color=I("black"), shape=21) +
  scale_x_continuous(lim = c(0, quantile(diamonds$carat, 0.99)) ) +
  scale_y_continuous(lim = c(0, quantile(diamonds$price, 0.99)) ) +
  ggtitle("Diamonds: Price vs. Carat") + 
  geom_smooth(method = 'lm') +
  stat_poly_eq(aes(label = paste0("atop(", ..eq.label.., ",", ..rr.label.., ")")), 
               formula = formula1, 
               parse = TRUE)
pricecarat
```


## Now create a linear model of price as a function of carat and color.
### What difference does it make if you convert color to a character vector before you make the model? Note that this is probably what you want to do in general, unless you really know what you’re doing.
### Make some plots of the raw data, and of the model fits, to make an argument as to which model is more useful. 
    
```{r q2, include=TRUE}
#linear model of price as a function of carat and color
carat_price_color <- lm(price ~ carat + color, data = diamonds)
carat_price_color
summary(carat_price_color)

#a linear model of price as a function of carat and color
ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1) +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = "Color", 
                                          reverse = F,
                                          override.aes = list(alpha = 1, 
                                                              size = 2))) +  
  ggtitle('Price of Diamonds as a Function of Carat and Color') + labs(x = "Carat", y = "Price ($)") + 
  stat_poly_eq(aes(label = paste0("atop(", ..eq.label.., ",", ..rr.label.., ")")), 
               formula = formula1, 
               parse = TRUE) +
    geom_smooth(method = 'lm') +
  facet_wrap(vars(color))
```

# Objects and Methods

```{r q3, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r q4, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```